---
authors:
- admin
date: "2019-09-21"
layout: post
title: "Tracking NVDA's share price"
subtitle: Nvidia's stock price compared to it's competitors.
summary: Nvidia's stock was outperforming it's industry peers during 2017 - 2019. I analyse their performance. 
tags: [fundamentals, finance, yahoo-finance]
comments: true
output:
  html_document:
    code_folding: hide
---

```{r setup, include = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(kableExtra)
library(tidyquant)
library(reshape2)
library(directlabels)
```

I had been tracking the performance of Nvidia for a while since they were reporting impressive earnings and share price growth over recent years. I conducted a study (DCF analysis) to see if Nvidia was fairly priced at this time. I had it marked down as a \$141 fair value per share based on the DCF I constructed, however it was trading between a high of \$280 to a low of \$130 from September and to December of 2018.

Let's download some fundamental data for Nvida and compare it's performance to it's industry peers.


```{r, include = TRUE, message = FALSE, warning = FALSE}

# A function to collect fundamental data from Yahoo finance (Income Statement, Balance Sheet and Cash Flow statements, just skip past this part.)

getFin <- function(stock){
  if ("rvest" %in% installed.packages()) {
    library(rvest)
    }else{
      install.packages("rvest")
      library(rvest)
      }
  for (i in 1:length(stock)) {
    tryCatch(
      {
        # Collect the Income Statement data
        link <- "https://finance.yahoo.com/quote/"
        link <- paste0(link, stock[i], "/financials?p=", stock[i])
        wahis.session <- html_session(link)
        p <- wahis.session %>%
          html_nodes(xpath = '//*[@id="Col1-1-Financials-Proxy"]/section/div[3]/table')%>%
          html_table(fill = TRUE)
        IncomeStatement <- p[[1]]
        colnames(IncomeStatement) <- paste(IncomeStatement[1,])
        IncomeStatement <- IncomeStatement[-c(1,5,12,20,25),]
        names_row <- paste(IncomeStatement[,1])
        IncomeStatement <- IncomeStatement[,-1]
        IncomeStatement <- apply(IncomeStatement, 2, function(x){gsub(",","",x)})
        IncomeStatement <- as.data.frame(apply(IncomeStatement, 2, as.numeric))
        rownames(IncomeStatement) <- paste(names_row)
        temp1 <- IncomeStatement
        
        # Collect the Balance Sheet data
        link <- "https://finance.yahoo.com/quote/"
        link <- paste0(link, stock[i],"/balance-sheet?p=", stock[i])
        wahis.session <- html_session(link)
        p <- wahis.session %>%
          html_nodes(xpath = '//*[@id="Col1-1-Financials-Proxy"]/section/div[3]/table')%>%
          html_table(fill = TRUE)
        BalanceSheet <- p[[1]]
        colnames(BalanceSheet) <- BalanceSheet[1,]
        BalanceSheet <- BalanceSheet[-c(1,2,17,28),]
        names_row <- BalanceSheet[,1]
        BalanceSheet <- BalanceSheet[,-1]
        BalanceSheet <- apply(BalanceSheet, 2, function(x){gsub(",","",x)})
        BalanceSheet <- as.data.frame(apply(BalanceSheet, 2, as.numeric))
        rownames(BalanceSheet) <- paste(names_row)
        temp2 <- BalanceSheet
        
        # Collect the Cash Flow data
        link <- "https://finance.yahoo.com/quote/"
        link <- paste0(link, stock[i], "/cash-flow?p=", stock[i])
        wahis.session <- html_session(link)
        p <- wahis.session %>%
          html_nodes(xpath = '//*[@id="Col1-1-Financials-Proxy"]/section/div[3]/table')%>%
          html_table(fill = TRUE)
        CashFlow <- p[[1]]
        colnames(CashFlow) <- CashFlow[1,]
        CashFlow <- CashFlow[-c(1,3,11,16),]
        names_row <- CashFlow[,1]
        CashFlow <- CashFlow[,-1]
        CashFlow <- apply(CashFlow, 2, function(x){gsub(",","",x)})
        CashFlow <- as.data.frame(apply(CashFlow, 2, as.numeric))
        rownames(CashFlow) <- paste(names_row)
        temp3 <- CashFlow
        
        assign(paste0(stock[i],'.f'),value = list(IncomeStatement = temp1, BalanceSheet = temp2, CashFlow = temp3), envir = parent.frame())
        },
      error = function(cond){
        message(stock[i], "Give error ",cond)
        }
      )
  }
}

```

We can inspect Nvidia's recent performance by scraping Yahoo finance. The original `getFinancials` option from the package `quantmod` became defunct a little while ago, however I have been using this function without problems ever since. You can just add tickers to the `symbols` object to download more company financial accounts.

```{r, include = TRUE, message = FALSE, warning = FALSE}
symbols <- c("NVDA")
getFin(symbols)
symbols.f <- sapply(symbols, function(x) { paste0(x, ".f") })
tickers <- list2env(mget(symbols.f))

IS <- lapply(tickers, "[[", "IncomeStatement")
BS <- lapply(tickers, "[[", "BalanceSheet")
CF <- lapply(tickers, "[[", "CashFlow")

IS <- as.data.frame(IS)
BS <- as.data.frame(BS)
CF <- as.data.frame(CF)
```

### **Nvidia's Income Statement data.**

```{r, include = TRUE, message = FALSE, warning = FALSE}

IS %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

### **Nvidia's Balance Sheet Information.**

```{r, include = TRUE, message = FALSE, warning = FALSE}
BS %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

### **Nvidia's Cash Flow Statements**

```{r, include = TRUE, message = FALSE, warning = FALSE}

CF %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

Nvidia's fundamentals looked solid after some analysis, but taking into account per share ratios, I changed my opinion. Their share price seemed overvalued compared to their business performance.

I wanted to track their performance and compare it to Nvidia's competitors. I downloaded price data using the `tidyquant` package. 

```{r, include = TRUE, message = FALSE, warning = FALSE}

############### Collect the stock price and benchmark data #############

nvda <- "NVDA" %>%
  tq_get(get = "stock.prices",
         from =  "2017-01-01",
         to = Sys.Date())

amd <- "AMD" %>%
  tq_get(get = "stock.prices",
         from =  "2017-01-01",
         to = Sys.Date())

txn <- "TXN" %>%
  tq_get(get = "stock.prices",
         from =  "2017-01-01",
         to = Sys.Date())

intc <- "INTC" %>%
  tq_get(get = "stock.prices",
         from =  "2017-01-01",
         to = Sys.Date())

nasdaq <- "^IXIC" %>%
  tq_get(get = "stock.prices",
         from = "2017-01-01",
         to = Sys.Date())
```

Joining all the stock price data together along with a benchmark NASDAQ:

### **Stock Price Data**

```{r, include = TRUE, message = FALSE, warning = FALSE}
########## Join the data together ###########
# Ra is the asset price
# Rb is the benchmark price (NASDAQ)
RaRb <- nvda %>%
  full_join(., amd, by = "date") %>%
  full_join(., intc, by = "date") %>%
  full_join(., nasdaq, by = "date") %>%
  select(date, adjusted.x, adjusted.y, adjusted.x.x, adjusted.y.y) %>%
  setNames(c("date", "NVDA", "AMD", "INTC", "NASDAQ"))

RaRb %>%
  head(6) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

It's still quite difficult to compare Nvidia's performance based on the *raw* stock price data, so I normalise the data.

```{r, include = TRUE, message = FALSE, warning = FALSE}
############# Normalise the data ################## 
normalise_series <- function(xdat) xdat / coredata(xdat)[1]

normalised_RaRb <- RaRb %>%
  mutate(NVDA = normalise_series(NVDA),
         AMD = normalise_series(AMD),
         INTC = normalise_series(INTC),
         NASDAQ = normalise_series(NASDAQ))
```

#### **The first few observations normalised**
```{r, include = TRUE, message = FALSE, warning = FALSE}
normalised_RaRb %>%
  head() %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
  
```

#### **The last few observations normalised**
```{r, include = TRUE, message = FALSE, warning = FALSE}
normalised_RaRb %>%
  tail() %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
  
```

The data looks a lot more comparable. I plot the results in `ggplot2`

```{r, include = TRUE, message = FALSE, warning = FALSE}
normalised_RaRb %>%
  melt(id = "date") %>%
  ggplot(aes(x = date, y = value, color = variable)) +
  geom_line(aes(linetype = variable)) +
  #scale_color_manual(values=c("red", "black", "darkblue", "yellow", "green")) +
  labs(title = "Normalised NVDA, AMD, INTC and NASDAQ Benchmark",
       x = "", subtitle = "From January 2017 - to Sys.Date()", y = "Normalised Price US$", color = "") +
  scale_y_continuous(labels = scales::dollar) +
  theme_tq(base_size = 8) +
  geom_dl(aes(label = variable), method = list(dl.combine("last.points"), cex = 1)) +
  scale_color_manual(values = c(NASDAQ = 'black',
                                NVDA = 'darkgreen',
                                AMD = 'red',
                                INTC = 'blue')) +
  theme(legend.position = "none")
```

So \$1 invested in Nvidia back in Jan 2017 would have given you aprox \$2.75 return if you sold at the back end of 2018, however the price today would be around \$1.60. Compare that with AMD you would have got the same return of \$2.75 if you sold at the same time as Nvidia but today your return would be more or less the same at aprox \$2.75 despite a drop in AMD's share price at the same time as NVDA.